---
title: "Data Analysis with R - Titanic"
output:
  pdf_document: default
  html_notebook: default
---
```{r}



# In this project I am goint to visualize Titanic dataset in R.


```


## Import the Libraries
```{r}
library(ggplot2) # This library for graphs
library(ggthemes) # This one is for graphs 's themes 
library(dplyr)  # This one is for combine train and test datasets


```


## Loading Data
```{r}
train <- read.csv("train.csv", stringsAsFactors = F)
test  <- read.csv("test.csv", stringsAsFactors = F)

options( warn = -1 )


```


## Combine data sets
bind_rows : Needing to combine datasets with non-matching columns

```{r}
full_data  <- dplyr::bind_rows(train,test)


```


Details of our dataset
```{r}
str(full_data)


```

Summary of our columns
```{r}
summary(full_data)

```


## Exploratory Analysis


### Age vs Survived
- Here I am going to compare persons age with if they are survived or not 
- The reason for use like this: "full_data[1:891,]", because I just want to analyze train part
```{r}
ggplot(full_data[1:891,], aes(Age, fill =factor(Survived))) + # aes= aesthetic mapping 
  geom_histogram(bins = 30) +
  theme_economist() +
  xlab("Age") + 
  scale_fill_discrete(name = "Survived") +
  ggtitle("Age vs Survived")

```




### Sex vs Survived

Here I am going to compare persons gender with if they are survived or not 
```{r}
ggplot(full_data[1:891,], aes(Sex, fill = factor(Survived))) + 
  geom_bar(stat = "count", position = 'dodge') + 
  theme_economist() +
  xlab("Sex") +
  ylab("Count") +
  scale_fill_discrete(name = "Survived") +
  ggtitle("Sex vs Survived")


```



### Age vs Sex vs Survived 

Here I am going to compare persons age and gender with if they are survived or not 

```{r}
ggplot(full_data[1:891,], aes(Age, fill = factor(Survived))) + 
  geom_histogram(bins = 30) + 
  theme_economist_white() +
  xlab("Age") +
  ylab("Count") +
  facet_grid(.~Sex) +
  scale_fill_discrete(name = "Survived") +
  ggtitle(" Age vs Sex vs Survived ")


```




### Pclass vs Survived

Here I am going to compare persons platform class with if they are survived or not 

```{r}
ggplot(full_data[1:891, ], aes(Pclass, fill = factor(Survived))) +
  geom_bar(stat = "count") +
  theme_economist() +
  xlab("Pclass") +
  ylab("Count") +
  scale_fill_discrete(name = "Survived") +
  ggtitle(" Pclass vs Survived")


```


### Pclass vs Sex vs Age vs Survived

Here I am going to compare persons platform class, gender, and age with if they are survived or not 

```{r}
ggplot(full_data[1:891,], aes(x = Age, y = Sex)) +
  geom_jitter(aes(colour = factor(Survived))) +
  theme_few() +
  theme(legend.title = element_blank()) +
  facet_wrap(~Pclass) +
  labs(x = "Age", y="Sex", title = "Pclass vs Sex vs Age vs Survived") +
  scale_fill_discrete(name = "Survived") +
  scale_x_continuous(name = "Age", limits = c(0,81))


```



Here as you can see, names labeled (for instance; Mr, Ms, etc.)

```{r}
head(as.character(full_data$Name))


```



Now I am going to show them on a table

```{r}
# Extract titles
full_data$Title <- gsub('(.*, )|(\\..*)', '', full_data$Name)

# Titles by Sex
table(full_data$Sex, full_data$Title)


```



As you can realize from the table some persons are labeled as Mlle, Mme, etc. Here I am going to replace them with Miss and Mrs. Also I am going to reassign officer and royalty which is used rarely.

```{r}
# Reassign rare titles
officer <- c('Capt', 'Col', 'Don', 'Dr', 'Major', 'Rev')
royalty <- c('Dona', 'Lady', 'the Countess', 'Sir', 'Jonkheer')

# Reassign Mlle, Ms, Mme and officer, and royalty
full_data$Title[full_data$Title == 'Mlle']    <- 'Miss'
full_data$Title[full_data$Title == 'Ms']      <- 'Miss'
full_data$Title[full_data$Title == 'Mme']     <- 'Mrs'

full_data$Title[full_data$Title %in% royalty]    <- 'Royalty'
full_data$Title[full_data$Title %in% officer]    <- 'Officer'

# Here I am going to seperate Surname from Name
full_data$Surname <- sapply(full_data$Name, function(x) strsplit(x, split = '[,.]')[[1]][1])


```



Here I am going to compare persons title with if they are survived or not 

```{r}
ggplot(full_data[1:891,], aes(Title, fill = factor(Survived))) +
  geom_bar(stat = "count") +
  theme_few() +
  xlab('Title') +
  ylab('Count') +
  scale_fill_discrete(name = "Survived") +
  ggtitle("Title vs Survived") 

```
















